\ZpL: a p-adic precision package

### 2018-02-03 ZpL/ZpL.tex
Reorganization of examples (all in Section 2)

If provably correct results are desired, interval arithmetic provides an alternative to floating point.  Instead of just tracking an approximation to the answer, the package also tracks a radius within which the true result lies.  This method is commonly used for $p$-adic computations since the ultrametric property of $p$-adic fields frequently keeps the radius small.  Computations remain fairly efficient with this approach, but numerical instability can still lead to dramatic losses in precision, (see Section \ref{SOMOS-4} for example).

@@

\section{Examples}

@@

@@

In linear algebra, characteristic polynomials are notoriously hard to 
compute: see \cite{caruso-roe-vaccon:15, caruso-roe-vaccon:17}. We 
illustrate this with the following example (using a division-free 
algorithm for the computation of the characteristic polynomial) :

@@

Our package can be applied to computation with $p$-adic polynomials. A 
natural example is that of the computation of GCD, whose stability has 
been studied in \cite{caruso:2017}.

A naive implementation of the Euclidean algorithm
can produce different behaviour depending 
on the type of implementation of the field
of $p$-adic coefficients.

@@

\section{Presentation of the package}

@@

\section{More examples}

\subsection{SOMOS-4}

The SOMOS-4 sequence is a well-known example
of a computation with a large gap between
the theoretical loss in precision
and that of a naive computation (see \cite{caruso-roe-vaccon:14a}).
We show here that precision can be saved
even when using a generic implementation
of the SOMOS iteration.

\subsection{Matrix multiplication}
The multiplication of many matrices has been studied
in \cite{caruso-roe-vaccon:15} and is known to often 
provide an heavy diffused digits phenomenon.
Here is a simple example.
% The result for \ZpFP has not been
% added, but its first digits coincide with that of \ZpLC.

\subsection{Gröbner bases}
Our package can be applied 
on complex computations like that of
Gröbner bases using generic
Gröbner bases algorithms.

As we can see, some loss in precision occurs in the
Buchberger algorithm and is avoided thanks to \ZpL.
%Had we used \ZpFP, many padded zeroes or one would have appeared.

>>>>

If provably correct results are desired, interval arithmetic provides an alternative to floating point.  Instead of just tracking an approximation to the answer, the package also tracks a radius within which the true result lies.  This method is commonly used for $p$-adic computations since the ultrametric property of $p$-adic fields frequently keeps the radius small.  Computations remain fairly efficient with this approach, but numerical instability can still lead to dramatic losses in precision, (see Section \ref{sec:demo} for many examples).

@@

\section{Short demonstration}
\label{sec:demo}

@@

\noindent \textbf{SOMOS 4.}
%
A first example is the SOMOS-4 sequence. It is defined by the recurrence: 
$$u_{n+4} = \frac{u_{n+1} u_{n+3} + u_{n+2}^2}{u_n}$$
and is known for its high numerical instability (see 
\cite{caruso-roe-vaccon:14a}). Nevertheless, the \ZpL package saves 
precision even when using a generic unstable implementation of the SOMOS 
iteration.

@@

Many generic algorithms of linear algebra lead to quite important 
instability when they are used with $p$-adic numbers. In many cases, 
our package \ZpL rubs this instability without having to change the 
algorithm, nor the implementation.

\noindent \textbf{Matrix multiplication.}
%
As revealed in \cite{caruso-roe-vaccon:15}, a first simple example where 
this phenomenon appears is simply matrix multiplication.
This might be surprising because no division occurs in this situation.
Observe nevertheless the difference between \ZpCR and \ZpLC.

\noindent
On the aforementioned example, we notice that \ZpCR is unable to decide 
whether the product vanishes or not. Having good estimates on the 
precision is therefore very important in such situations.

\noindent \textbf{Characteristic polynomials.}
%
Characteristic polynomials are notoriously hard to compute 
\cite{caruso-roe-vaccon:15, caruso-roe-vaccon:17}. We illustrate this 
with the following example (using the default algorithm of \sage for the 
computation of the characteristic polynomial, which is a division free 
algorithm in this setting) :

@@

Our package can be applied to computation with $p$-adic polynomials.

\noindent \textbf{Euclidean algorithm.}
%
A natural example is that of the computation of GCD, whose stability has 
been studied in \cite{caruso:2017}.
A naive implementation of the Euclidean algorithm can produce different 
behaviour depending on the type of implementation of the field of 
$p$-adic coefficients.

@@

\noindent \textbf{Gröbner bases.}
%
Our package can be applied on complex computations like that of Gröbner 
bases using generic Gröbner bases algorithms.

\noindent
As we can see, some loss in precision occurs in the
Buchberger algorithm and is avoided thanks to \ZpL.
%Had we used \ZpFP, many padded zeroes or one would have appeared.

\section{What's behind the scene?}

@@
